---
title: "data_analysis_v2"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r load_packages}
# # Ensure that pacman is installed for package management and loading.
# if (!require("pacman")) install.packages("pacman")
# # for data reading wrangling and visualization
pacman::p_load(tidyverse) 
# for working directories
pacman::p_load(here) 
# # for cross tabulation and data cleaning
# pacman::p_load(janitor) 
# for working with strings
pacman::p_load(glue) 
# For randomized inference, also loads randomizr and estimatr 
pacman::p_load(ri2) 
# for marginal effects from lineal regressions
pacman::p_load(margins)
# Tests for linear regression models
pacman::p_load(lmtest)
pacman::p_load(car)
# Tables
pacman::p_load(kableExtra)
# for updated ggplot2 theme
pacman::p_load(hrbrthemes)
# for updated ggplot2 colorblind-friendly scheme
pacman::p_load(ggthemes)
# theme_set(hrbrthemes::theme_ipsum())
pacman::p_load(reshape2)
# for plotting of covariate balance
pacman::p_load(cobalt)
# for matching only
pacman::p_load(MatchIt)
library(stargazer)
library(usmap) 
library(ggplot2)  
library(maps)  
#install.packages("mapproj") 
library(mapproj)
#install.packages("viridis")
library(viridis)
library(dplyr)
library(ggthemes) 
library(sjPlot)
library(sjmisc)

```

```{r read_data}
#download the data from GitHub
data <-'https://raw.githubusercontent.com/gsbDBI/ALP301-spr21-project3/main/rla_clean_5_12.csv'
df <- read.csv(data, strip.white = TRUE)
rm(data) #remove data csv file

```

```{r treatment table}
#table() creates a contingency table of counts of observations at each combination of treat_pseudo and treat_real
with(df, table(treatment_group, useNA = 'ifany')) %>%  # "ifany" includes the NA values in the table
  knitr::kable() %>%                                            #kabel(x, format) generates tables 
  # add in a header to label what we're cross-tabulating with
  add_header_above(c('treat_group' = 2)) %>% #add_header_above(x, col_name=col_span)
  kableExtra::kable_styling(bootstrap_options = "striped") #additional styling options

```

```{r attention check}
#see who is in the top 2.5% of least amount of time spent on the survey 
quantile(df$duration_sec, c(0.025, 0.95, .975)) 
# we see the duration is 88 seconds for the top 2.5% and 97.5% (those who spent the least amount)

# make a density plot of the distribution of total time spent on the survey between 
df %>%
  filter( duration_sec < 1050, duration_sec > 88 ) %>%
  ggplot( aes(x=duration_sec)) +
    geom_density(fill="#69b3a2", color="#e9ecef", alpha=0.8)
```

```{r graph of income}
# make bar graph of income
p2 <- ggplot(data = subset(df, income != ""), aes(income)) + 
  geom_bar(binwidth=3, fill="#1c3ca8cf", color="#e9ecef", alpha=0.9) +
  ggtitle("Income of respondents") +
theme(
  plot.background = element_rect(fill = "white"),
  panel.background = element_rect(fill = "white"),
  axis.line.x = element_line(color = "grey")
  ) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1), plot.title = element_text(size=10))

#ggsave("income.png", plot = p2)

p3 <- ggplot(data = df, aes(treatment_group)) +
  geom_bar(binwidth=3, fill="#1c3ca8cf", color="#e9ecef", alpha=0.9) +
  ggtitle("Treatment groups of respondents") +
theme(
  plot.background = element_rect(fill = "white"),
  panel.background = element_rect(fill = "white"),
  axis.line.x = element_line(color = "grey")
  ) +
  xlab("Treatment Group") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1), plot.title = element_text(size=10))

#ggsave("treatment_group.png", plot = p3)
```

```{r cov_redefine}
#included in cleaning script
# df$gender_female <- ifelse(df$gender == "Female" | df$gender == "Transgender Female", 1, 0)
# 
# df$parent_yes <- ifelse(df$parent == "Yes", 1, 0)
# 
# df$race_hispanic <- ifelse(grepl("Latino or race_hispanic", df$race), 1, 0)
# 
df$edu_4college <- ifelse(df$edu == "4-year college degree" | df$edu == "Postgraduate degree (MA, MBA, JD, PhD, etc.)", 1, 0)
# df$edu_hs <- ifelse(df$edu == "High school graduate (high school diploma or equivalent including GED)", 1, 0)

```

```{r map of respondents}
# load US state map data
us_states <- map_data("state")  
head(us_states)  

# aggregate number of respondents per state from your survey data
bystate <- df %>% 
  group_by(state) %>% 
  dplyr::summarise(
    numbperstate = n()
  )

# merge RLA data w/maps data

bystate$state <- tolower(bystate$state)

# fix errors
bystate[bystate$state=="new hampshir",]$state <- "new hampshire"
bystate[bystate$state=="minnesotta",]$state <- "minnesota"
bystate[bystate$state=="west virgina",]$state <- "west virginia"

us_states_df <- us_states %>% 
  left_join(bystate, by = c("region" = "state"))

p <- ggplot(data = us_states_df,
             mapping = aes(x = long, y = lat,
                           group = group, fill = numbperstate))+
  geom_polygon(color = "gray90", size = 0.1) +
  coord_map(projection = "albers", lat0 = 39, lat1 = 45) +
  scale_fill_continuous(type = "viridis", name = "Number of respondents \n per state") +
  theme_map() +
  theme(legend.position="bottom",
        axis.line=element_blank(),
        axis.text=element_blank(),
        axis.ticks=element_blank(),
        axis.title=element_blank(),
        panel.background=element_blank(),
        panel.border=element_blank(),
        panel.grid=element_blank(),
        text = element_text(size=14))

#ggsave("map.png", plot = p)

```


```{r balance_plots}
df$age <- as.numeric(df$age)
# exclude data with missing covariates
no_na_df <- df %>% drop_na('age', 'state', 'income', 'party', 'gender_female', 'parent_yes', 'race_hispanic', 'edu_4college', 'edu_hs')

love.plot(treatment_group_num ~ age + gender_female + parent_yes + state + income_clean + race_hispanic + edu_4college + edu_hs + dv_pre_state_conf + dv_pre_national_conf + party,
                 data = no_na_df, 
          binary = "std", 
          limits = c(-.5, .5),
          thresholds = c(m = .1)) # threshold determines the dotted lines on the graph

m.out <- matchit(dummy_treat ~ age + gender_female + parent_yes + state + income_clean + race_hispanic + edu_4college + edu_hs + dv_pre_state_conf + dv_pre_national_conf + party,
                 data = no_na_df, 
                method = "nearest", replace = TRUE) # perform matching
# plot covariate balance with and without matching
love.plot(m.out, binary = "std", thresholds = c(m = .1))

```
Interpretation: treatment-covariate correlations is very low.

```{r compare tests with attention check failures and 2SD of duration}
# people within 2 SD of duration
sd2 = sd(df$duration_sec)*2

duration_sd2_df <- no_na_df %>% 
  filter(abs(duration_sec - mean(duration_sec)) < sd2)

# only removes 15 people so would not sig affect results
rm(duration_sd2_df)

# people who passed attention check
atten_check_df <- no_na_df %>% 
  filter(attention_pass == 1)

love.plot(treatment_group_num ~ age + gender_female + parent_yes + state + income_clean + race_hispanic + edu_4college + edu_hs + dv_pre_state_conf + dv_pre_national_conf + party,
                 data = atten_check_df, 
          binary = "std", 
          limits = c(-.5, .5),
          thresholds = c(m = .1)) # threshold determines the dotted lines on the graph

m.out <- matchit(dummy_treat ~ age + gender_female + parent_yes + state + income_clean + race_hispanic + edu_4college + edu_hs + dv_pre_state_conf + dv_pre_national_conf + party,
                 data = atten_check_df, 
                method = "nearest", replace = TRUE) # perform matching
# plot covariate balance with and without matching
love.plot(m.out, binary = "std", thresholds = c(m = .1))

# results look pretty similar for those who failed and did not fail the attention check
```

```{r region_probs}
# Check whether treatment assignment probabilities vary by region
# lm_robust performs linear regression with robust standard errors
# including a -1 in the regression model drops the intercept
# ~ indicates the regression equation: dependent variable ~ independent variables
estimatr::lm_robust(dummy_treat ~ state - 1, data = df) 
```

Another way of checking for balance is to run a multivariate regression of our treatment assignment variable on covariates. With this method, instead of a t-test, we use an F-test for the hypothesis that all coefficients (non-strata) are equal to zero. If a coefficient on a covariate is not close to zero, that would mean that assignment was not balanced on that covariate.

```{r lm_ftest_balance}
# regression of covariates on treatment assignment variable
balance_lm <- lm(treatment_group_num ~ age + gender_female + parent_yes + state + income_clean + race_hispanic + edu_4college + edu_hs + party + dv_pre_state_conf + dv_pre_national_conf, data = no_na_df) # factor(region) encodes the string variable as a factor for analysis


# Test whether all coefficients from the balancce_lm regression are equal to zero
# using heteroskedasticity-robust standard errors, denoted by hc2
car::linearHypothesis(balance_lm, c("gender_female = 0", "age = 0", "parent_yes = 0", "race_hispanic = 0", "edu_hs = 0", "dv_pre_state_conf = 0", "dv_pre_national_conf = 0"),
  test = "F", white.adjust = "hc2", singular.ok = TRUE)

summary(balance_lm)
```

None of the coefficients is significant, meaning that there is no significantly different pre-experiment covariates that determine treatment status.

```{r balance_table}
# Define a function to generate a balance check table using two-sample t-test 
# Parameters: "cov_list" includes covariates, "treat" indicates the treatment, "alpha" is the significance level
t_table <- function(data, cov_list, treat,
                    alpha = 0.05) {
  # for each covariate in cov_list, apply a function to conduct the t-test
  out <- lapply(cov_list, function(x) {
      tres <- t.test(data[[x]] ~ data[[treat]]) # t-test result between control and treatment groups' covariate x
    # capture the group means for both control and treatment, and the p-value
    c(mean_control = as.numeric(tres$estimate[1]), mean_treat = as.numeric(tres$estimate[2]), p_value = tres$p.value)    
  })
  
  # save results by binding the results for all above covariates in cov_list by row (rbind)
  out <- as.data.frame(do.call(rbind, out))
  # combine covaraite names and results by column (cbind)
  out <- cbind(covariate = cov_list, out)
  # get ride of non-important strings in the names
  names(out) <- gsub("[^0-9A-Za-z_]", "", names(out)) # gsub(pattern, replacement, x) replaces pattern in x with replacement
  
  # code presentation of p-value according to whether they are below a threshold, for e.g., 0.001
  out$p_value <- ifelse(out$p_value < 0.001,
    "<0.001",
    round(out$p_value, 3) # rounding of p values to 3 decimal places
  )
  return(out)
}

# Apply above t-t_table function to our dataset and selected covariates
balance_table <- t_table(
  data = df,
  c("gender_female", "parent_yes", "race_hispanic", "edu_hs", "trust_federal", "trust_state"),
  "dummy_treat"
)

# output the balance table
balance_table %>% 
  knitr::kable() %>%
  kableExtra::kable_styling(bootstrap_options = "striped")
```

We see that we reject the null hypothesis of equal means for all of the covariates, as all the P-values are > 0.01.

```{r region_balance}
# balance table for Democrats
balance_dems <- t_table(
  data = subset(df, party=="Democrat"), # subsetting data to those with "party" equal to "Democrat"
  c("gender_female", "parent_yes", "race_hispanic", "edu_hs", "trust_federal", "trust_state"),
  "dummy_treat"
)

# balance table for Republicans
balance_repubs <- t_table(
  data = subset(df, party=="Republican"), # subsetting data to those with "party" equal to "Republican"
  c("gender_female", "parent_yes", "race_hispanic", "edu_hs", "trust_federal", "trust_state"),
  "dummy_treat"
)

balance_indpt_other <- t_table(
  data = subset(df, party=="Independent" | party == "Other Party"), # subsetting data to those with "party" equal to "Republican"
  c("gender_female", "parent_yes", "race_hispanic", "edu_hs", "trust_federal", "trust_state"),
  "dummy_treat"
)

# combine the 4 t-tables, excluding first column of covariate names for the latter 3, so that only one covariate name column is there
balance_party = cbind(balance_dems, balance_repubs %>% select(2:4), balance_indpt_other %>% select(2:4))

# Format: adding in headers
x <- knitr::kable(balance_party, digits = 2) %>% kable_styling()
# add in a header to label what we're cross-tabulating with
add_header_above(x, c('', 'Democrats'=3, 'Republicans'=3, 'Other and Independent'=3))


```

```{r dem vs repub trust}
# of those who identify as democrats how trust 2016 and 2020 

# p4 <- ggplot(data = df, aes(accuracy_2016)) +
#   geom_bar(binwidth=3, fill="#1c3ca8cf", color="#e9ecef", alpha=0.9) +
#   ggtitle("Treatment groups of respondents") +
# theme(
#   plot.background = element_rect(fill = "white"),
#   panel.background = element_rect(fill = "white"),
#   axis.line.x = element_line(color = "grey")
#   ) +
#   xlab("2016 accuracy") +
#   theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1), plot.title = element_text(size=10))

```

```{r basic regression with dummy treat}
df[c(2,6:10,12,19:21,23,27,29,30,32,35:42,44,45)] <- lapply(df[c(2,6:10,12,19:21,23,27,29,30,32,35:42,44,45)], factor)

# run regressions to estimate treatment effect with robust standard errors

# run the regression for state confidence level 
reg1 <- lm(dv_post_state_conf ~ dummy_treat + dv_pre_state_conf, data = df)

# run the regression for national confidence level
reg2 <- lm(dv_post_national_conf ~ dummy_treat + dv_pre_national_conf, data = df)

# run the regression with controls for state level
reg3 <- lm(dv_post_state_conf ~ dummy_treat + dv_pre_state_conf + age + gender_female + parent_yes + state + income_clean + race_hispanic + race_white + race_black + race_asian + edu_4college + edu_hs + party, data = df)

# run the regression with controls for national level
reg4 <- lm(dv_post_national_conf ~ dummy_treat + dv_pre_national_conf + age + gender_female + parent_yes + state + income_clean + race_hispanic + race_white + race_black + race_asian + edu_4college + edu_hs + party, data = df)

```

```{r regression with treatment groups}
# run the regression with treatment groups
df$treatment_group <- relevel(factor(df$treatment_group), ref = "Control")

#state level with treatment groups
reg5 <- lm(dv_post_state_conf ~ treatment_group + dv_pre_state_conf + age + gender_female + parent_yes + state + income_clean + race_hispanic + race_white + race_black + race_asian + edu_4college + edu_hs + party, data = df)

# national with treatment groups
reg6 <- lm(dv_post_national_conf ~ treatment_group + dv_pre_national_conf + age + gender_female + parent_yes + state + income_clean + race_hispanic + race_white + race_black + race_asian + edu_4college + edu_hs + party, data = df)

# transform estimates to dataframe and drop unnecessary rows
# lm_treatment1_te <- bind_rows(list(tidy(reg5))) %>% 
#   mutate(term = gsub('treatment_group', '', term)) %>%
#   filter((term %in% c("(Intercept)", "dv_pre_state_conf", "Income", "edu", "state")))
# 
# # plot treatment effect
# ggplot(lm_treatment1_te, aes(x = term,y =estimate)) +
#   geom_point(position=position_dodge(width=0.5)) +
#   geom_errorbar(aes(x=term,ymin = estimate - 1.96 * std.error,ymax = estimate + 1.96 * std.error),width = .1,position=position_dodge(width=0.5), color = "firebrick") + 
#   xlab('Treatment') +
#   ylab('Estimate') +
#   theme(plot.title = element_text(hjust = 0.5),
#         axis.text.x = element_text(angle = 45, vjust = 1, hjust =1)) + 
#   geom_hline(yintercept = 0, colour = 'grey60', linetype = 2) +
#   coord_cartesian(ylim=c(-.5, 0.5)) + 
#   ggtitle('Treatment1 Post-Test Treatment Effect (conditioned on pre-test)')
```

```{r regression table, echo=FALSE, message=TRUE, warning=FALSE, results='asis'}
stargazer(reg1, reg2, reg3, reg4, 
          omit = "state", "income_clean",
          column.labels=c("state","national", "state with controls", "national with controls"),
          title = "Basic Regression Results for Treatment Effect")

```

```{r coefficient plot control vs treated}
library(jtools)

plot_coefs(
  reg1, reg2, reg3, reg4, 
  coefs = c("Treatment Effect \n vs. Control \n (95% C.I.)" = "dummy_treat"),
  ci_level = 0.95,
  model.names = c("State", "National", "State w/ controls", "National w/ controls"),
  scale = TRUE)
```

```{r coefficient plot of treatment effects}
plot_coefs(
  reg5, reg6,
  coefs = c("Bipartisan Auditors" = "treatment_groupBipartisan", "Handcount" = "treatment_groupHandcount", "Local Officials" = "treatment_groupLocal", "Loser" = "treatment_groupLoser", "RLA percentage" = "treatment_groupRL_percentage", "Soup \n 95% C.I." = "treatment_groupSoup"),
  ci_level = 0.95,
  model.names = c("State", "National"),
  x.label = "Estimate",
  scale = TRUE)

```

```{r compare_age_dist}
# df %>% 
#   ggplot(aes(x = age, # aes(x, y, ...) defines how variables are mapped into the aesthetics of the plot 
#             color = treatment_group,
#             fill = treatment_group
#              )
#          ) +
#   # Density plot
#   geom_density(alpha = 0.3) +    # alpha controls the transparency
#   # use a colorblind friendly color palette
#   scale_color_colorblind() +
#   scale_fill_colorblind() +
#   theme(legend.position = "top") # legend position

```

```{r compare_gender_dist}
# df %>% 
#   ggplot(aes(x = gender, # aes(x, y, ...) defines how variables are mapped into the aesthetics of the plot 
#             color = treatment_group,
#             fill = treatment_group
#              )
#          ) +
#   # Density plot
#   geom_density(alpha = 0.3) +    # alpha controls the transparency
#   # use a colorblind friendly color palette
#   scale_color_colorblind() +
#   scale_fill_colorblind() +
#   theme(legend.position = "top") # legend position

```

```{r compare trust in government}
# plot trust in federal government
# df %>% 
#   ggplot(aes(x = trust_federal, # aes(x, y, ...) defines how variables are mapped into the aesthetics of the plot 
#             color = treatment_group,
#             fill = treatment_group
#              )
#          ) +
#   # Density plot
#   geom_density(alpha = 0.3) +    # alpha controls the transparency
#   # use a colorblind friendly color palette
#   scale_color_colorblind() +
#   scale_fill_colorblind() +
#   theme(legend.position = "top") # legend position
# 
# # trust in state government
# df %>% 
#   ggplot(aes(x = trust_state, # aes(x, y, ...) defines how variables are mapped into the aesthetics of the plot 
#             color = treatment_group,
#             fill = treatment_group
#              )
#          ) +
#   # Density plot
#   geom_density(alpha = 0.3) +    # alpha controls the transparency
#   # use a colorblind friendly color palette
#   scale_color_colorblind() +
#   scale_fill_colorblind() +
#   theme(legend.position = "top") # legend position
```

```{r difference_in_means_simple}
# simple difference in means for those treated with a message
with(df, mean(dv_post_state_conf[dummy_treat == 1]) - mean(dv_post_state_conf[dummy_treat == 0])) #with(data, expr, …) evaluates the expr on the data

with(df, mean(dv_post_national_conf[dummy_treat == 1]) - mean(dv_post_national_conf[dummy_treat == 0])) #with(data, expr, …) evaluates the expr on the data
```

We use this `difference_in_means` function because it allows us to subset data, and conveniently provides the standard error and other statistics for the estimate. 

```{r difference_in_means}
# Use the estimatr package for difference in means estimates
estimatr::difference_in_means(dv_state_treatment_diff ~ dummy_treat, data = df) 
# difference_in_means(y ~ x, data) computes mean(y when x==1) - mean(y when x==0) in data, along with standard errors and p-values from two-sided t-tests

estimatr::difference_in_means(dv_national_treatment_diff ~ dummy_treat, data = df) 

```

```{r difference_in_means_subsample}
# difference in means for women at state level
estimatr::difference_in_means(dv_state_treatment_diff ~ dummy_treat, data = df, subset = gender_female == "1")

# difference in means for women at national level
estimatr::difference_in_means(dv_national_treatment_diff ~ dummy_treat, data = df, subset = gender_female == "1")

# difference in means for age at state level
# age < 40
estimatr::difference_in_means(dv_state_treatment_diff ~ dummy_treat, data = df, subset = age < 40)
# age > 40
estimatr::difference_in_means(dv_state_treatment_diff ~ dummy_treat, data = df, subset = age > 40)

# difference in means for age at national level
# age > 40
estimatr::difference_in_means(dv_national_treatment_diff ~ dummy_treat, data = df, subset = age > 40)

# difference in means for HS education at state level
estimatr::difference_in_means(dv_state_treatment_diff ~ dummy_treat, data = df, subset = edu_hs == "1")

# difference in means for HS education at national level
estimatr::difference_in_means(dv_national_treatment_diff ~ dummy_treat, data = df, subset = edu_hs== "1")

# difference in means for college education at state level
estimatr::difference_in_means(dv_state_treatment_diff ~ dummy_treat, data = df, subset = edu_4college == "1")

# difference in means for college education at national level
estimatr::difference_in_means(dv_national_treatment_diff ~ dummy_treat, data = df, subset = edu_4college == "1")

```

```{r gender interaction}
# GENDER

#state level interaction for gender
reg_state_fem <- lm(dv_post_state_conf ~ treatment_group*gender_female + dv_pre_state_conf + age + parent_yes + state + income_clean + race_hispanic + race_white + race_black + race_asian + edu_4college + edu_hs + party, data = df)

plot_model(reg_state_fem, type = "pred", terms = c("gender_female","treatment_group"))
plot_model(reg_state_fem, type = "int", terms = c("gender_female","treatment_group"))

# national level interaction for gender
reg_natl_fem <- lm(dv_post_national_conf ~ treatment_group*gender_female + dv_pre_national_conf + age + parent_yes + state + income_clean + race_hispanic + race_white + race_black + race_asian + edu_4college + edu_hs + party, data = df)

plot_model(reg_natl_fem, type = "pred", terms = c("gender_female", "treatment_group"))
plot_model(reg_natl_fem, type = "int", terms = c("treatment_group", "gender_female"))
```

```{r age interaction}
# AGE 

#state level 
reg_state_age <- lm(dv_post_state_conf ~ treatment_group*age + gender_female + dv_pre_state_conf + parent_yes + state + income_clean + race_hispanic + race_white + race_black + race_asian + edu_4college + edu_hs + party, data = df)

summary(reg_state_age)

plot_model(reg_state_age, type = "pred", terms = "age")


# national level 
reg_natl_age <- lm(dv_post_national_conf ~ treatment_group*age + gender_female + dv_pre_national_conf + parent_yes + state + income_clean + race_hispanic + race_white + race_black + race_asian + edu_4college + edu_hs + party, data = df)

summary(reg_natl_age)

plot_model(reg_natl_age, type = "pred", terms = "age")
plot_model(reg_natl_age, type = "pred", terms = c("age", "party"))

plot_model(reg_natl_age, type = "int", terms = c("treatment_group", "age"))
plot_model(reg_natl_age, type = "pred", terms = c("treatment_group", "age"))
```

```{r educ interaction}
# education

# state only HS 
reg_state_edu_hs <- lm(dv_post_state_conf ~ treatment_group*edu_hs + dv_pre_state_conf + age + gender_female + parent_yes + state + income_clean + race_hispanic + race_white + race_black + race_asian + edu_4college + party, data = df)

plot_model(reg_state_edu_hs, type = "int", terms = c("treatment_group", "edu_4college"))
plot_model(reg_state_edu_hs, type = "pred", terms = c("edu_4college","treatment_group"))

# national only HS
reg_natl_edu_hs <- lm(dv_post_national_conf ~ treatment_group*edu_hs + dv_pre_national_conf + age + gender_female + parent_yes + state + income_clean + race_hispanic + race_white + race_black + race_asian + edu_4college + party, data = df)

plot_model(reg_natl_edu_hs, type = "int", terms = c("treatment_group", "edu_4college"))
plot_model(reg_natl_edu_hs, type = "pred", terms = c("edu_4college","treatment_group"))

# state college or above
reg_state_edu <- lm(dv_post_state_conf ~ treatment_group*edu_4college + dv_pre_state_conf + age + gender_female + parent_yes + state + income_clean + race_hispanic + race_white + race_black + race_asian + edu_hs + party, data = df)

plot_model(reg_state_edu, type = "int", terms = c("treatment_group", "edu_4college"))
plot_model(reg_state_edu, type = "pred", terms = c("edu_4college","treatment_group"))

# national college or above
reg_natl_edu <- lm(dv_post_national_conf ~ treatment_group*edu_4college + dv_pre_national_conf + age + gender_female + parent_yes + state + income_clean + race_hispanic + race_white + race_black + race_asian + edu_hs + party, data = df)

plot_model(reg_natl_edu, type = "int", terms = c("treatment_group", "edu_4college"))
plot_model(reg_natl_edu, type = "pred", terms = c("edu_4college","treatment_group"))

```

```{r income interaction}

# income - figure out better way to visualize

# state 
reg_state_inc <- lm(dv_post_state_conf ~ treatment_group*income_clean +  dv_pre_state_conf + age + gender_female + parent_yes + state + race_hispanic + race_white + race_black + race_asian + edu_4college + edu_hs + party, data = df)

summary(reg_state_inc)

plot_model(reg_state_inc, type = "int", terms = c("treatment_group", "income_clean"))
plot_model(reg_state_inc, type = "pred", terms = c("income_clean", "treatment_group"))

# national
reg_natl_inc <- lm(dv_post_national_conf ~ treatment_group*income_clean + dv_pre_national_conf + age + gender_female + parent_yes + state + race_hispanic + race_white + race_black + race_asian + edu_4college + edu_hs + party, data = df)

summary(reg_natl_inc)

plot_model(reg_natl_inc, type = "int", terms = c("treatment_group", "income_clean"))
plot_model(reg_natl_inc, type = "pred", terms = c("income_clean", "treatment_group"))
```

```{r always_never believer and political party}
# always vs never believer
# df$always_believer <- to_factor(df$always_believer)
# df$never_believer <- to_factor(df$never_believer)
# state never believer
reg_state_never <- lm(dv_post_state_conf ~ treatment_group*never_believer + dv_pre_state_conf + age + gender_female + parent_yes + state + income_clean + race_hispanic + race_white + race_black + race_asian + edu_4college + edu_hs + party, data = df)

summary(reg_state_never)

plot_model(reg_state_never, type = "int", terms = c("treatment_group", "never_believer"))
plot_model(reg_state_never, type = "pred", terms = c("never_believer", "treatment_group"))

# national never believer
reg_natl_never <- lm(dv_post_national_conf ~ treatment_group*never_believer + dv_pre_national_conf + age + gender_female + parent_yes + state + income_clean + race_hispanic + race_white + race_black + race_asian + edu_4college + edu_hs + party, data = df)

summary(reg_natl_never)

plot_model(reg_natl_never, type = "int", terms = c("treatment_group", "never_believer"))
plot_model(reg_natl_never, type = "pred", terms = c("treatment_group", "never_believer"))

# state always believer
reg_state_always <- lm(dv_post_state_conf ~ treatment_group*always_believer + dv_pre_state_conf + age + gender_female + parent_yes + state + income_clean + race_hispanic + race_white + race_black + race_asian + edu_4college + edu_hs + party, data = df)

plot_model(reg_state_always, type = "int", terms = c("treatment_group", "always_believer"))
plot_model(reg_state_always, type = "pred", terms = c("always_believer", "treatment_group"))

# national always believer
reg_natl_always <- lm(dv_post_national_conf ~ treatment_group*always_believer + dv_pre_national_conf + age + gender_female + parent_yes + state + income_clean + race_hispanic + race_white + race_black + race_asian + edu_4college + edu_hs + party, data = df)

plot_model(reg_natl_always, type = "int", terms = c("treatment_group", "always_believer"))
plot_model(reg_natl_always, type = "pred", terms = c("always_believer", "treatment_group"))

# political party 

# state level
df$party <- to_factor(df$party)

reg_state_party <- lm(dv_post_state_conf ~ treatment_group*party + dv_pre_state_conf + age + gender_female + parent_yes + state + income_clean + race_hispanic + race_white + race_black + race_asian + edu_4college + edu_hs, data = df)

plot_model(reg_state_party, type = "int", terms = c("treatment_group", "party"))
plot_model(reg_state_party, type = "pred", terms = c("party", "treatment_group"))

# national level
reg_natl_party <- lm(dv_post_national_conf ~ treatment_group*party +  dv_pre_national_conf + age + gender_female + parent_yes + state + income_clean + race_hispanic + race_white + race_black + race_asian + edu_4college + edu_hs, data = df)

plot_model(reg_natl_party, type = "int", terms = c("treatment_group", "party"))
plot_model(reg_natl_party, type = "pred", terms = c("party", "treatment_group"))

# race 
# state level race_hispanic Y/N

reg_state_race <- lm(dv_post_state_conf ~ treatment_group*race_hispanic + dv_pre_state_conf + age + gender_female + parent_yes + state + income_clean +  + race_white + race_black + race_asian + edu_4college + edu_hs + party, data = df)

summary(reg_state_race)

plot_model(reg_state_race, type = "int", terms = c("treatment_group", "race_hispanic"))
plot_model(reg_state_race, type = "pred", terms = c("race_hispanic", "treatment_group"))

# national level race_hispanic Y/N
reg_natl_race <- lm(dv_post_national_conf ~ treatment_group*race_hispanic + dv_pre_national_conf + age + gender_female + parent_yes + state + income_clean +  + race_white + race_black + race_asian + edu_4college + edu_hs + party, data = df)

summary(reg_natl_race)

plot_model(reg_natl_race, type = "int", terms = c("treatment_group", "race_hispanic"))
plot_model(reg_natl_race, type = "pred", terms = c("race_hispanic", "treatment_group"))


```

```{r filter out extreme views pre test}
reg_middle_views_state <- lm(dv_post_state_conf ~ treatment_group + dv_pre_state_conf + age + gender_female + parent_yes + state + income_clean + race_hispanic + race_white + race_black + race_asian + edu_4college + edu_hs + party, data = df, subset = (dv_pre_state_conf < 10 & dv_pre_state_conf > 1))

summary(reg_middle_views_state)

plot_model(reg_middle_views_state, type = "pred", terms = c("treatment_group"))

reg_middle_views_natl <- lm(dv_post_national_conf ~ treatment_group + dv_pre_national_conf + age + gender_female + parent_yes + state + income_clean + race_hispanic + race_white + race_black + race_asian + edu_4college + edu_hs + party, data = df, subset = (dv_pre_state_conf < 10 & dv_pre_state_conf > 1))

summary(reg_middle_views_natl)

```

```{r plot control vs. treatment results}
df$treatment_group <- relevel(factor(df$treatment_group), ref = "Control")


```
